Automated Topic Naming

### 2011-08-29 ese/ese.tex
removed ctrl-m line endings from abram - I think we can tell the server to ignore these. Also fixed small bug where hash sign was used

%

\RequirePackage{fix-cm}

%

%\documentclass{svjour3}                     % onecolumn (standard format)

%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)

\documentclass[smallextended]{svjour3}       % onecolumn (second format)

%\documentclass[twocolumn]{svjour3}          % twocolumn

%

\smartqed  % flush right qed marks, e.g. at end of proof

%

\usepackage{graphicx}

 \usepackage{booktabs}

\usepackage[caption=false]{subfig}

 \usepackage[colorinlistoftodos,backgroundcolor=yellow]{todonotes} %disable

 \usepackage{color}

@@

\newcommand{\XXX}[1]{\textcolor{red}{{\it \textbf{[XXX: #1]}}}}

\hyphenation{Max-DB data-base}

\usepackage{comment}

\newtheorem{mydef}{definition}

\newcommand{\slbl}[1]{\textbf{\textsf{#1}}}

\usepackage{multirow}%,multicolumn}

\usepackage{slashbox}

\usepackage{wrapfig}

\usepackage{booktabs, colortbl}

\usepackage{amsmath}

\usepackage{url}

\usepackage[pdfstartview=FitH,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},

\journalname{Empirical Software Engineering}

@@

\title{Automated Topic Naming

%\thanks{Grants or other notes

%about the article that should go on the front page should be

%placed here. General acknowledgments should be placed at the end of the article.}

}

\subtitle{Supporting Cross-project Analysis of Software Maintenance Activities}

%\titlerunning{Short form of title}        % if too long for running head

@@

\author{

Abram Hindle \and

Neil A. Ernst \and

Michael W. Godfrey \and

John Mylopoulos

}

%\authorrunning{Short form of author list} % if too long for running head

\institute{

  Abram Hindle \at

  Dept. of Computer Science\\

  University of Alberta\\

  Edmonton, AB, CANADA\\

  \email{abram@softwareprocess.es}

\and

  Neil A. Ernst \at

  Dept. of Computer Science\\

  University of Toronto\\

  Toronto, Ontario, CANADA\\

  \email{nernst@cs.toronto.edu}

\and

  Michael W. Godfrey \at

  David Cheriton School of Computer Science\\

  University of Waterloo\\

  Waterloo, Ontario, CANADA\\

  \email{migod@uwaterloo.ca}

\and

  John Mylopoulos \at

  Dept. Information Eng. and Computer Science\\

  University of Trento\\

  Trento, ITALY\\

  \email{jm@disi.unitn.it}

}

\date{Received: date / Accepted: date}

% The correct dates will be entered by the editor

@@

\begin{abstract}

% %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known

Researchers have employed a variety of techniques to extract underlying

topics that relate to software development artifacts.  Typically, these

techniques use semi-unsupervised machine-learning algorithms to suggest

candidate word-lists.  However, word-lists are

difficult to

interpret in the absence of meaningful summary labels.  Current topic modeling

techniques assume manual labelling and do not use domain-specific knowledge

to improve, contextualize, or describe results for the developers.

We propose a solution: \emph{automated labelled topic extraction}.

  Topics are extracted

using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered

from source control systems such as CVS and BitKeeper.

These topics are given labels from a generalizable

cross-project taxonomy, consisting of non-functional

requirements. Our approach was evaluated with experiments and case studies

on three large-scale RDBMS projects: MySQL, PostgreSQL and MaxDB.  The case studies

show that labelled topic extraction can produce appropriate,

context-sensitive labels relevant to these projects, which provides fresh

insight into their evolving software development activities.

%160 words

\keywords{ Software maintenance \and Repository mining \and Latent dirichlet allocation \and Topic models}

@@

\section{Introduction}

A key problem for practising software maintainers is gaining an

understanding of \emph{why} a system has evolved the way it has.

This is different than \emph{how} a system has evolved, because the behaviour of change itself is the

\emph{how}.

Looking back on streams of artifacts scattered across different

repositories, inferring what activities were performed, when, and for

what reasons, is hard to do without expert advice from the developers

involved.

%new stuff

% edit this. awkward

In this work we seek to provide a method of automatically labelling  development topics extracted from commit logs.

Topic modeling is a machine learning technique which creates

multinomial distributions of words extracted from a text corpus.

This technique infers the hidden structure of a corpus using posterior

inference: the probability of the hidden structure given the data.

Topic models are useful in software maintenance because they summarize

the key concepts in a corpus, such as source code, commit comments, or

mailing-list messages, by identifying statistically co-occurring words.

Among other uses, it can quickly give developers an overview of where significant

activity has occurred, and gives managers or maintainers a sense of

project history.

%A major challenge for topic modeling is to accurately identify the

%summary label for a topic; in the previous case, perhaps

%\emph{``event-handling''} would be appropriate.

%Current approaches are manual: they rely on human expertise to

%recognize the appropriate label.

%Furthermore, to date the results of topic modeling are project

%specific, and not generalizable.

%This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.

%migod

While machine learning techniques can automatically identify clumps of

commonly recurring terms, devising an appropriate summary label for

each clump/topic is harder.

A given topic extracted from a set of commit logs might consist of the following terms: \emph{ ``listener change remove add fire''}.

This topic might reasonably be labelled as

``\emph{event handling}'' by a developer who understands the domain well,

despite the fact that this label does not appear in the word list itself.

Current approaches to topic labelling rely on manual intervention by

human experts, and also are limited to project-specific topic labels.

In this paper, we introduce \emph{labelled topic extraction}, an

approach to topic labelling that creates labels automatically that are

project independent.

%is both automatic and creates labels that are project independent.

% Changes to software systems, revisions, do not occur without

% reason.

% The reason or purpose is often related to requirements

% Often one revision might have multiple purposes ranging from

% maintenance issues, quality issues, requirements, features or a

% response to a bug report.

% These purposes often are exhibited as latent topics that relate to

% many development artifacts such as:

%  version control revisions, bug reports, and mailing-list discussions.

%In our previous work we dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}.

%We observed that topic trends were often non-functional requirements (NFRs).

%NFRs have the property of being cross-domain and widely applicable.

%In this sense, they are useful abstractions for developer conversations about different software projects.

% harp on data mining more

%In general, the mining of software artifacts tends to be very project specific, yet NFRs are not.

% I was told if you're going to use a paren, you don't need to say it.

%There is a series of standards on NFRs, such as ISO9126~\cite{iso9126}, specifically intended to apply to projects of varying types.

%These standards indicate our approach is possible. %, if perhaps not universally accepted).

% NFRs are prevalent in almost all software systems.

% In this paper we will show that NFR related topics occur quite frequently.

%XXXXXX this sentence is confusing

%This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering,

%specifically NFRs.

%In particular, we use software quality models to generalize topics across projects.

%This allows us to mine software systems with the intention of comparing their NFR-related development topics.

% in order to achieve generalizable, cross project data mining results.

%%%% THIS PARAGRAPH ARGUES ABOUT THE CONCRETE APPLICATION OF

%%%% TOPIC MODELLING AND LABELLED TOPIC EXTRACTION

%%%% IT NEEDS TO BE TIGHTER BUT IT ALSO NEEDS TO BE HERE

% AH: Topics take a long to read

% AH: Project Dashboards, summarization, software process recovery

%migod

In general, the fruits of mining software artifacts are often project

specific and hard to generalize.

However, in our previous work we

investigated \emph{topic trends} --- that is, topics that recur over

time --- we observed that topic trends often corresponded to

non-functional requirements (NFRs)~\cite{Hindle09ICSM}, which is

further emphasized in this paper due to the large numbers of NFR

labelled topics.

This is encouraging, as NFRs have the property of being cross-domain

and widely applicable.

In this sense, they are useful abstractions for developer

conversations about different software projects.

Furthermore, there is a series of standards on NFRs, such as ISO9126 \cite{iso9126}, that are specifically intended to apply to projects of varying

types; this suggests that our goal of trying to extract NFR-related development topics, such as those related to software quality models, holds promise.

Concrete applications of topics and \emph{labelled topic extraction}

range from project dashboards to annotating software artifacts such as revisions and bug reports with

NFR-related tags.

Project dashboard~\cite{dashboard} are typically employed by managers and are used to

provide quick summaries of the effort put into a software project. In this case,

labelled topic extraction would allow managers to track effort

related to NFR topics, such as portability.

These techniques also allow for the annotation of commit comments

 and other software artifacts with NFRs.

This would enable querying of bug reports and artifacts by relevant NFRs.

For instance a manager can confirm if their developers were

focused on \emph{usability} by looking for

 \emph{usability}-relevant revisions and bug reports.

In this paper, we describe \emph{automated labelled topic extraction}. It addresses two gaps in the topic mining literature:

\begin{enumerate}

  \item Topic mining of software has been limited to one project at a time.

This is because traditional topic mining techniques are specific to a particular data-set.

\textit{Automated labelled topic extraction} allows for comparisons \textit{between} projects.

  \item Topic modeling creates word lists that require interpretation by the user to assign meaning.

Like (1), this means

that it is difficult to discuss results independent of the project context.

Our technique automatically, or with some initial training, assigns labels across projects.

This paper makes the following contributions:

\begin{itemize}

\item introduces the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels;

%\item evaluates an unsupervised method of automated topic labelling (word-lists);

%%\item Our unsupervised method can be applied to projects without training;

%%\item

%evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);

%%\item

%evaluates a supervised method of labelling topics with multiple NFRs

%(multi-label machine learning);

\item evaluates three kinds of automatic topic labelling methods:

  semi-unsupervised labelling  of topics (word-lists), supervised labelling of

  topics with a single NFR (machine learning),

  and supervised labelling of topics with multiple NFRs (multi-label

  machine learning);

%  of automated topic labelling (word-lists);

%\item Our unsupervised method can be applied to projects without training;

%\item

%evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);

%\item

%evaluates a supervised method of labelling topics with multiple NFRs (multi-label machine learning);

\item examines how NFRs correlate with the work of individual developers;

\item provides a method of cross-project analysis via topic labelling; and

 applies these techniques to visualize NFRs over time, and to analyze maintenance activities.

@@

%We first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.

We begin by discussing related work in Section \ref{sec:related}.

Next, we describe how we generated our data (Section \ref{sec:wordlist}). For semi-unsupervised classification (Section \ref{sec:unsuplabelling}), we

begin by creating word-lists to signify when a topic matches an NFR label. We then apply our classifier and analyze the results.

%and how we created the word-lists used in unsupervised classification.

%We show that labels with their topics can be learned and used to classify other data-sets, either without training

%Our goal is to label a given topic with either an NFR from our taxonomy,

 %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.

In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.

%We then present visualizations of named topics and their trends over time to aid communication and analysis.

To demonstrate an application of labelled topic extraction, we use an exploratory case study of two open source database systems to show how named

topics can be compared between projects  (Section \ref{sec:analysis}).

The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and future work.

\section{Previous Work}

\label{sec:related}

The idea of extracting higher-level \emph{concerns} and \emph{topics}, also known as

 \emph{concepts}, \emph{aspects} or \emph{requirements},

has been approached from documentation-based and repository-based

perspectives.

Cleland-Huang and her colleagues have investigated mining requirements documents for non-functional requirements (NFR) (quality

requirements)~\cite{Cleland-Huang2006}.

Their approach is similar to ours, as they mined keywords from NFR

catalogues.

They differ because they mine requirements documents where as we mine revisions.

%One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}.

They demonstrated a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs.

Instead, they developed a supervised classifier by using human experts to identify an NFR training set.

Our research differs because we use a more comprehensive set of terms based on the taxonomy we chose.

Another difference is that we make cross-project comparisons instead of focusing

on a single project.

%They relied on requirements documents instead of version control

%histories that we use.

%The objective of Cleland-Huang's study was to identify new NFRs for system development,

%yet our objective was to recover those latent NFRs from commit-log messages of the project.

%There are several reasons we did not follow this route.

%One, we believe we have a more comprehensive set of terms due to the taxonomy we chose.

%Secondly, we wanted to compare across projects.

%Their technique was not compared across different projects and the applicability of the training set to different corpora is unclear.

%A common taxonomy allows us to make cross-project comparison.

%Although subject to the assumption that all projects conceive of these terms in the same way.

%Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development)

%our study assumes these NFRs are latent in the commit-log messages of the project.

Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system.

Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words,

including common related forms (similar to word stemming), meronymy and synonymy.

They used WordNet for word roots as they felt the synonyms would be

non-specific and cause errors.

%, problems we also encountered.

Mockus et al. validated their labels with system developers.

Since we study multiple projects, instead of a single project, these

kind of interviews are somewhat infeasible (particularly in the distributed world of open-source software).

Another approach is to extract concerns from software repositories.

Marcus et al.~\cite{marcus04wcre} use Latent Semantic Indexing (LSI)

to identify commonly occurring concerns for software maintenance.

The

concerns are given by the user, and LSI is used to retrieve them from

a corpus.

Topic modelling generates topics that are independent of a user query, and relate only to word frequencies in the corpus.

%Some results were interesting, but their precision was quite low.

With ConcernLines, Treude et al.~\cite{treude09cl} show tag occurrence

using colour and intensity.

They mine

developer created tags

%change request tags (such as `milestone 3') and used these

in order to analyze the evolution of a single product.

% make evolutionary analyses of a single product.

The presence of a well-maintained set of tags is obviously essential to the success of this technique.

%XXXXXXXXXXXX OPTIONAL

% Not sure we need this

%Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.

% They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess:

% ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p. 388]{mens08icsm}''.

%This paper examines the notions of quality in terms of a consistent ontology, as Menset al. call for in their conclusions.

% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)

In Baldi et al.~\cite{Baldi2008}, topics are named manually: human

experts read the highest-frequency members of a topic and assign a

label accordingly.

%XXX Possible to cut this

As discussed earlier, given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}.

The labels are reasonable enough, but still require an expert in the field to determine them.

Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called \emph{`Oracle'}

in the MySQL case, since Oracle owns MySQL.

%We do this differently.

%This paper does this differently.

Our approach differs: first of all, we automate the process of naming the topics; secondly, we label topics with project-independent terms, in order

to permit cross-project comparison.

Mei et al.~\cite{Mei2007} use context information to automatically name topics.

They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase.

They do not use external domain-specific information as we do, but we

do not generate phrases from the topics.

Finally, in Ernst et al.~\cite{ernst10refsq}, we describe our earlier

project, similar to this, that identifies changes in quality requirements

in GNOME software projects; this approach was more exploratory, had

less validation, uses different word-lists, solely uses text-matching,

and does not leverage machine learning strategies.

Hindle et al.~\cite{Hindle09ICSM} propose a windowed method of topic

analysis that we extend with labelled

topics, NFRs and new visualizations.

%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software.

% this reduces the line length and drops a ref even though we still have a scacchi kickin around

%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software.

%Their work discusses the source of the requirements and how they are used in the development process.

% German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes.

%None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.

\todo{re-add this related work?}

% XXX vanity or space..

%XXXXXXXXXXXX OPTIONAL

% NE 0113 this seems too early to mention 'result'.

% Hindle et al. \cite{Hindle2007} examined release patterns in OSS. They showed that there is a difference between projects regarding maintenance

% techniques. This supports our result that software qualities are not discussed with the same frequency across projects.

\section{Study Design and Execution}

Figure \ref{fig:process} gives an outline of our methodology.

We began by gathering source data and creating topic models. For semi-unsupervised labelling, we generated three sets of word-lists as signifiers for NFRs.

%e created labels using a well-known NFR taxonomy.

In supervised learning, we trained our data with manual annotations in order to match topics with NFRs. Finally, these topics were used to analyze the

role of NFRs in software maintenance.

\begin{figure}

  \centering

>>>>

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
 \usepackage{booktabs}
\usepackage[caption=false]{subfig}
 \usepackage[colorinlistoftodos,backgroundcolor=yellow]{todonotes} %disable
 \usepackage{color}

@@

\newcommand{\XXX}[1]{\textcolor{red}{{\it \textbf{[XXX: #1]}}}}
\hyphenation{Max-DB data-base}

\usepackage{comment}
\newtheorem{mydef}{definition}
\newcommand{\slbl}[1]{\textbf{\textsf{#1}}}
\usepackage{multirow}%,multicolumn}
\usepackage{slashbox}
\usepackage{wrapfig}
\usepackage{booktabs, colortbl}
\usepackage{amsmath}
\usepackage{url}
\usepackage[pdfstartview=FitH,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
\journalname{Empirical Software Engineering}

@@

\title{Automated Topic Naming}
%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}

\subtitle{Supporting Cross-project Analysis of Software Maintenance Activities}
%\titlerunning{Short form of title}        % if too long for running head

@@

\author{
Abram Hindle \and
Neil A. Ernst \and
Michael W. Godfrey \and
John Mylopoulos
}
%\authorrunning{Short form of author list} % if too long for running head

\institute{
  Abram Hindle \at 
  Dept. of Computer Science\\
  University of Alberta\\
  Edmonton, AB, CANADA\\
  \email{abram@softwareprocess.es}
\and
  Neil A. Ernst \at
  Dept. of Computer Science\\
  University of British Columbia\\
  Vancouver, BC, CANADA\\
  \email{nernst@cs.ubc.ca}
\and
  Michael W. Godfrey \at
  David Cheriton School of Computer Science\\
  University of Waterloo\\
  Waterloo, Ontario, CANADA\\
  \email{migod@uwaterloo.ca}
\and 
  John Mylopoulos \at
  Dept. Information Eng. and Computer Science\\
  University of Trento\\
  Trento, ITALY\\
  \email{jm@disi.unitn.it}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

@@

\begin{abstract}
% %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
Researchers have employed a variety of techniques to extract underlying
topics that relate to software development artifacts.  Typically, these
techniques use semi-unsupervised machine-learning algorithms to suggest
candidate word-lists.  However, word-lists are
difficult to 
interpret in the absence of meaningful summary labels.  Current topic modeling 
techniques assume manual labelling and do not use domain-specific knowledge
to improve, contextualize, or describe results for the developers.  
We propose a solution: \emph{automated labelled topic extraction}.
  Topics are extracted 
using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
from source control systems such as CVS and BitKeeper.  
These topics are given labels from a generalizable
cross-project taxonomy, consisting of non-functional
requirements. Our approach was evaluated with experiments and case studies
on three large-scale RDBMS projects: MySQL, PostgreSQL and MaxDB.  The case studies
show that labelled topic extraction can produce appropriate,
context-sensitive labels relevant to these projects, which provides fresh
insight into their evolving software development activities.
%160 words
\keywords{ Software maintenance \and Repository mining \and Latent dirichlet allocation \and Topic models}

@@

\section{Introduction}
A key problem for practising software maintainers is gaining an
understanding of \emph{why} a system has evolved the way it has. 
This is different than \emph{how} a system has evolved, because the behaviour of change itself is the
\emph{how}.
Looking back on streams of artifacts scattered across different
repositories, inferring what activities were performed, when, and for
what reasons, is hard to do without expert advice from the developers
involved. 
%new stuff
% edit this. awkward
In this work we seek to provide a method of automatically labelling  development topics extracted from commit logs.

Topic modeling is a machine learning technique which creates
multinomial distributions of words extracted from a text corpus. 
This technique infers the hidden structure of a corpus using posterior
inference: the probability of the hidden structure given the data. 
Topic models are useful in software maintenance because they summarize
the key concepts in a corpus, such as source code, commit comments, or
mailing-list messages, by identifying statistically co-occurring words. 
Among other uses, it can quickly give developers an overview of where significant
activity has occurred, and gives managers or maintainers a sense of
project history. 
%A major challenge for topic modeling is to accurately identify the
%summary label for a topic; in the previous case, perhaps
%\emph{``event-handling''} would be appropriate. 
%Current approaches are manual: they rely on human expertise to
%recognize the appropriate label. 
%Furthermore, to date the results of topic modeling are project
%specific, and not generalizable. 
%This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.

%migod
While machine learning techniques can automatically identify clumps of
commonly recurring terms, devising an appropriate summary label for
each clump/topic is harder.  
A given topic extracted from a set of commit logs might consist of the following terms: \emph{ ``listener change remove add fire''}. 
This topic might reasonably be labelled as
``\emph{event handling}'' by a developer who understands the domain well,
despite the fact that this label does not appear in the word list itself.  
Current approaches to topic labelling rely on manual intervention by
human experts, and also are limited to project-specific topic labels.  
In this paper, we introduce \emph{labelled topic extraction}, an
approach to topic labelling that creates labels automatically that are
project independent.
%is both automatic and creates labels that are project independent.

% Changes to software systems, revisions, do not occur without
% reason. 
% The reason or purpose is often related to requirements 
% Often one revision might have multiple purposes ranging from
% maintenance issues, quality issues, requirements, features or a
% response to a bug report.
% These purposes often are exhibited as latent topics that relate to
% many development artifacts such as:
%  version control revisions, bug reports, and mailing-list discussions.

%In our previous work we dealt with topic trends, which are topics that recur over time~\cite{Hindle09ICSM}. 
%We observed that topic trends were often non-functional requirements (NFRs). 
%NFRs have the property of being cross-domain and widely applicable. 
%In this sense, they are useful abstractions for developer conversations about different software projects.

% harp on data mining more
%In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. 
% I was told if you're going to use a paren, you don't need to say it.
%There is a series of standards on NFRs, such as ISO9126~\cite{iso9126}, specifically intended to apply to projects of varying types.
%These standards indicate our approach is possible. %, if perhaps not universally accepted).
% NFRs are prevalent in almost all software systems.
% In this paper we will show that NFR related topics occur quite frequently.
%XXXXXX this sentence is confusing
%This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering,
%specifically NFRs. 
%In particular, we use software quality models to generalize topics across projects. 
%This allows us to mine software systems with the intention of comparing their NFR-related development topics.
% in order to achieve generalizable, cross project data mining results.

%%%% THIS PARAGRAPH ARGUES ABOUT THE CONCRETE APPLICATION OF 
%%%% TOPIC MODELLING AND LABELLED TOPIC EXTRACTION
%%%% IT NEEDS TO BE TIGHTER BUT IT ALSO NEEDS TO BE HERE

% AH: Topics take a long to read
% AH: Project Dashboards, summarization, software process recovery

%migod
In general, the fruits of mining software artifacts are often project
specific and hard to generalize.  
However, in our previous work we
investigated \emph{topic trends} --- that is, topics that recur over
time --- we observed that topic trends often corresponded to
non-functional requirements (NFRs)~\cite{Hindle09ICSM}, which is
further emphasized in this paper due to the large numbers of NFR
labelled topics.  
This is encouraging, as NFRs have the property of being cross-domain
and widely applicable. 
In this sense, they are useful abstractions for developer
conversations about different software projects.  
Furthermore, there is a series of standards on NFRs, such as ISO9126 \cite{iso9126}, that are specifically intended to apply to projects of varying
types; this suggests that our goal of trying to extract NFR-related development topics, such as those related to software quality models, holds promise.

Concrete applications of topics and \emph{labelled topic extraction}
range from project dashboards to annotating software artifacts such as revisions and bug reports with
NFR-related tags.
Project dashboard~\cite{dashboard} are typically employed by managers and are used to
provide quick summaries of the effort put into a software project. In this case, 
labelled topic extraction would allow managers to track effort
related to NFR topics, such as portability.
These techniques also allow for the annotation of commit comments
 and other software artifacts with NFRs. 
This would enable querying of bug reports and artifacts by relevant NFRs. 
For instance a manager can confirm if their developers were
focused on \emph{usability} by looking for
 \emph{usability}-relevant revisions and bug reports.

In this paper, we describe \emph{automated labelled topic extraction}. It addresses two gaps in the topic mining literature:
\begin{enumerate}
  \item Topic mining of software has been limited to one project at a time. 
This is because traditional topic mining techniques are specific to a particular data-set. 
\textit{Automated labelled topic extraction} allows for comparisons \textit{between} projects. 
  \item Topic modeling creates word lists that require interpretation by the user to assign meaning. 
Like (1), this means
that it is difficult to discuss results independent of the project context. 
Our technique automatically, or with some initial training, assigns labels across projects.
\end{enumerate}

This paper makes the following contributions: 
\begin{itemize}
\item introduces the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels;

%\item evaluates an unsupervised method of automated topic labelling (word-lists);
%%\item Our unsupervised method can be applied to projects without training;
%%\item 
%evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
%%\item 
%evaluates a supervised method of labelling topics with multiple NFRs
%(multi-label machine learning);

\item evaluates three kinds of automatic topic labelling methods:
  semi-unsupervised labelling  of topics (word-lists), supervised labelling of
  topics with a single NFR (machine learning), 
  and supervised labelling of topics with multiple NFRs (multi-label
  machine learning);

%  of automated topic labelling (word-lists);
%\item Our unsupervised method can be applied to projects without training;
%\item 
%evaluates a supervised method of automatically labelling topics with a single NFR (machine learning);
%\item 
%evaluates a supervised method of labelling topics with multiple NFRs (multi-label machine learning);

\item examines how NFRs correlate with the work of individual developers;
\item provides a method of cross-project analysis via topic labelling; and
 applies these techniques to visualize NFRs over time, and to analyze maintenance activities.

@@

%We first introduces labelled topic extraction, and then uses that technique to analyze development activity in two large-scale examples.
We begin by discussing related work in Section \ref{sec:related}.
Next, we describe how we generated our data (Section \ref{sec:wordlist}). For semi-unsupervised classification (Section \ref{sec:unsuplabelling}), we
begin by creating word-lists to signify when a topic matches an NFR label. We then apply our classifier and analyze the results. 
%and how we created the word-lists used in unsupervised classification. 
%We show that labels with their topics can be learned and used to classify other data-sets, either without training 
%Our goal is to label a given topic with either an NFR from our taxonomy,
 %Section \ref{sec:unsuplabelling} presents our technique for using word-lists to do unsupervised classification, and describes the results.
In Section \ref{sec:suplabelling}, we manually annotate the topics, and use those annotations as training data for supervised classification.  
%We then present visualizations of named topics and their trends over time to aid communication and analysis. 
To demonstrate an application of labelled topic extraction, we use an exploratory case study of two open source database systems to show how named
topics can be compared between projects  (Section \ref{sec:analysis}). 
The paper concludes with a discussion of limitations (Section \ref{sec:limit}), and future work.

\section{Previous Work}
\label{sec:related}

The idea of extracting higher-level \emph{concerns} and \emph{topics}, also known as
 \emph{concepts}, \emph{aspects} or \emph{requirements},
has been approached from documentation-based and repository-based
perspectives.

Cleland-Huang and her colleagues have investigated mining requirements documents for non-functional requirements (NFR) (quality
requirements)~\cite{Cleland-Huang2006}. 
Their approach is similar to ours, as they mined keywords from NFR
catalogues.
They differ because they mine requirements documents where as we mine revisions.
%One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
They demonstrated a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
Our research differs because we use a more comprehensive set of terms based on the taxonomy we chose.
Another difference is that we make cross-project comparisons instead of focusing
on a single project.
%They relied on requirements documents instead of version control
%histories that we use.
%The objective of Cleland-Huang's study was to identify new NFRs for system development, 
%yet our objective was to recover those latent NFRs from commit-log messages of the project.

%There are several reasons we did not follow this route. 
%One, we believe we have a more comprehensive set of terms due to the taxonomy we chose. 
%Secondly, we wanted to compare across projects. 
%Their technique was not compared across different projects and the applicability of the training set to different corpora is unclear. 
%A common taxonomy allows us to make cross-project comparison. 
%Although subject to the assumption that all projects conceive of these terms in the same way.
%Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) 
%our study assumes these NFRs are latent in the commit-log messages of the project.

Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words,
including common related forms (similar to word stemming), meronymy and synonymy.
They used WordNet for word roots as they felt the synonyms would be
non-specific and cause errors.
%, problems we also encountered. 
Mockus et al. validated their labels with system developers.
Since we study multiple projects, instead of a single project, these
kind of interviews are somewhat infeasible (particularly in the distributed world of open-source software).

Another approach is to extract concerns from software repositories.
Marcus et al.~\cite{marcus04wcre} use Latent Semantic Indexing (LSI)
to identify commonly occurring concerns for software maintenance. 
The
concerns are given by the user, and LSI is used to retrieve them from
a corpus. 
Topic modelling generates topics that are independent of a user query, and relate only to word frequencies in the corpus.
%Some results were interesting, but their precision was quite low. 
With ConcernLines, Treude et al.~\cite{treude09cl} show tag occurrence
using colour and intensity. 
They mine
developer created tags 
%change request tags (such as `milestone 3') and used these 
in order to analyze the evolution of a single product.
% make evolutionary analyses of a single product. 
The presence of a well-maintained set of tags is obviously essential to the success of this technique.

%XXXXXXXXXXXX OPTIONAL
% Not sure we need this
%Mens et al.~\cite{mens08icsm} conducted an empirical study of Eclipse, a popular Java IDE, to verify the claims of Lehman~\cite{lehman97sms}.
% They concerned themselves with source code only, and found Law Seven, ``Declining Quality'', to be too difficult to assess:
% ``[we lacked an] appropriate measurement of the evolving quality of the system as perceived by the users \cite[p. 388]{mens08icsm}''. 
%This paper examines the notions of quality in terms of a consistent ontology, as Menset al. call for in their conclusions.

% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
In Baldi et al.~\cite{Baldi2008}, topics are named manually: human
experts read the highest-frequency members of a topic and assign a
label accordingly. 
%XXX Possible to cut this
As discussed earlier, given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
The labels are reasonable enough, but still require an expert in the field to determine them. 
Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called \emph{`Oracle'}
in the MySQL case, since Oracle owns MySQL. 
%We do this differently.
%This paper does this differently. 
Our approach differs: first of all, we automate the process of naming the topics; secondly, we label topics with project-independent terms, in order
to permit cross-project comparison.

Mei et al.~\cite{Mei2007} use context information to automatically name topics. 
They describe probabilistic labelling, using the frequency distribution of words in a topic to create a meaningful phrase. 
They do not use external domain-specific information as we do, but we
do not generate phrases from the topics.

Finally, in Ernst et al.~\cite{ernst10refsq}, we describe our earlier
project, similar to this, that identifies changes in quality requirements
in GNOME software projects; this approach was more exploratory, had
less validation, uses different word-lists, solely uses text-matching,
and does not leverage machine learning strategies. 
Hindle et al.~\cite{Hindle09ICSM} propose a windowed method of topic
analysis that we extend with labelled
topics, NFRs and new visualizations.

%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi02,scacchi05b} looked at the topic of requirements in open-source software. 
% this reduces the line length and drops a ref even though we still have a scacchi kickin around
%Massey~\cite{massey02icse} and Scacchi~\cite{scacchi05b} looked at the topic of requirements in open-source software. 
%Their work discusses the source of the requirements and how they are used in the development process. 
% German~\cite{german03gnome} looked at GNOME specifically, and listed several sources for requirements: leader interest, mimicry, brainstorming, and prototypes. 
%None of this work  addressed quality requirements in OSS, nor did it examine requirements trends.
\todo{re-add this related work?}
% XXX vanity or space.. 
%XXXXXXXXXXXX OPTIONAL
% NE 0113 this seems too early to mention 'result'.
% Hindle et al. \cite{Hindle2007} examined release patterns in OSS. They showed that there is a difference between projects regarding maintenance
% techniques. This supports our result that software qualities are not discussed with the same frequency across projects.

\section{Study Design and Execution}
Figure \ref{fig:process} gives an outline of our methodology.
We began by gathering source data and creating topic models. For semi-unsupervised labelling, we generated three sets of word-lists as signifiers for NFRs.
%e created labels using a well-known NFR taxonomy. 
In supervised learning, we trained our data with manual annotations in order to match topics with NFRs. Finally, these topics were used to analyze the
role of NFRs in software maintenance.
